---
title: "Configuration Notes"
output: html_document
date: "2022-11-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Notes

This is a file to help me collect configuration details about the various set-up processes necessary for the fungal/bacterial genome work addressed in this project.

### Conda - the Virtual Python Environment for BlobTools

I am using [Anaconda3](https://www.anaconda.com/products/distribution) (as of 11/09/22) to manage/access the required different packages and platforms.  Here is a brief [walkthrough](https://conda.io/projects/conda/en/stable/user-guide/getting-started.html) and a [cheat sheet](https://conda.io/projects/conda/en/stable/_downloads/cb0ffc4c7b1e6c0e716c066d2b077faf/conda-4.12.pdf) of important conda commands.

I made a directory in my Anaconda folder called "fungi", made to use Python 3.9.


### OpenSSL to download repos for conda

To pull BlobTools onto the computer, I need to add [OpenSSL](https://github.com/openssl/openssl), an encryption system for secure communication - if we don't have this tool on the computer, Conda can't pull BlobTools. 

To install, run this code while in Base:
```{r, eval=FALSE, echo=TRUE}
conda install -c anaconda openssl
```


### Blobtools Set-up

There are other versions of BlobTools that come packaged with some other softwareWe will be pulling Blobtools from [GitHub](https://github.com/DRL/blobtools)

In Anaconda terminal, ask the following commands: 

```{r, eval=FALSE, echo=TRUE}

conda create -n blobtools 
  # asks conda to make an environment called blobtools
conda activate blobtools
  # activates that environment

```

When I tried to install packages to the new blobtools environment, I found my Conda threw errors that OpenSSL was not available for the machine.  I tried the following fix to help (as suggested in the VERY long thread on [StackExchange](https://github.com/conda/conda/issues/11795):

```{r, eval=FALSE, echo=TRUE}
conda deactivate #goes back to (base)

conda update --all

# This is helpful in making sure all your packages are up-to-date, but doesn't ultimately help with this issue.
```
 
Note: This DID NOT work.  Instead, I had to install OpenSSL into blobtools from the (base) directory, as follows:
 
```{r, eval=FALSE, echo=TRUE}
conda install -n blobtools openssl
```
 
What a headache! Just to be sure, I installed all the other dependencies this way (from base, directly into blobtools). Probably not necessary.

I ran into a problem while trying to install "wget", though. This package is important for grabbing files via HTTPS (and other internet protocols) - iof we don't have this package, we won't be able to grab the NCBI databases required to BLAST our sequences against. When I try to do a normal install, Conda threw the following error: 

* PackagesNotFoundError: The following packages are not available from current channels: (and then it printed "wget"). 

THis happened regardless of remote install fromt he (base), or in the (blobtools) itself. I tried the following different installs to no avail:

```{r, eval=FALSE, echo=TRUE}
# These DON'T WORK
conda install -c conda-forge wget
conda install -c "conda-forge/label/cf202003" wget
```

There is a probably a simply fix that I'm misundertsanding but we'll figure it out. 

## Aside: BlobToolKit

There is also a tool kti that includes BlobTools2, Viewer,  [BlobToolKit](https://blobtoolkit.genomehubs.org/). Here is a [video tutorial](https://www.youtube.com/watch?v=SuKBKEH0cMA) from the package creators - note this install is for Mac/Linux systems. Before installing, you will need the following tools:

* [R](https://www.r-project.org/)
* [Git](https://git-scm.com/)

```{r, eval=FALSE, echo=TRUE}
# For install on Windows OS:
# Use Terminal to enter the following Git command.

git clone https://github.com/DRL/blobtools.git

```
